---
getwd(title: "ML project"
output: html_document
date: "2025-02-12"
---

```{r}
setwd("/Users/cindyyu/Desktop/CS4220")
library(readxl)
library(tidyverse)
library(caTools)
library(randomForest)
library(caret)
library(VIM)
library(ggplot2)
library(mice)
```

#READING IN FILTER PREDICTIONS, and MERGING THEM
```{r}
real1_df <- read.table("snv-parse-real1.txt", header = T)
real2.1_df <- read.table("snv-parse-real2_part1.txt", header = T)
syn1_df <- read.table("snv-parse-syn1.txt", header = T)
syn2_df <- read.table("snv-parse-syn2.txt", header = T)
syn3_df <- read.table("snv-parse-syn3.txt", header = T)
syn4_df <- read.table("snv-parse-syn4.txt", header = T)
syn5_df <- read.table("snv-parse-syn5.txt", header = T)

all_df <- rbind(real1_df, real2.1_df, syn1_df, syn2_df, syn3_df, syn4_df, syn5_df)
```

#READING IN TRUTHS, and MERGING THEM 
```{r}
#add headers: chromosome #, genomic start, genomic stop
real1_truth <- as.data.frame(read.table("/Users/cindyyu/Desktop/CS4220/Data/real1/real1_truth.bed",
                                header = FALSE, sep="\t",stringsAsFactors=FALSE, quote=""))
real2.1_truth <- as.data.frame(read.table("/Users/cindyyu/Desktop/CS4220/Data/real2_part1/real2_truth_chr1to5.bed",
                                header = FALSE, sep="\t",stringsAsFactors=FALSE, quote=""))
syn1_truth <- as.data.frame(read.table("/Users/cindyyu/Desktop/CS4220/Data/syn1/syn1_truth.bed",
                                header = FALSE, sep="\t",stringsAsFactors=FALSE, quote=""))
syn2_truth <- as.data.frame(read.table("/Users/cindyyu/Desktop/CS4220/Data/syn2/syn2_truth.bed",
                                header = FALSE, sep="\t",stringsAsFactors=FALSE, quote=""))
syn3_truth <- as.data.frame(read.table("/Users/cindyyu/Desktop/CS4220/Data/syn3/syn3_truth.bed",
                                header = FALSE, sep="\t",stringsAsFactors=FALSE, quote=""))
syn4_truth <- as.data.frame(read.table("/Users/cindyyu/Desktop/CS4220/Data/syn4/syn4_truth.bed",
                                header = FALSE, sep="\t",stringsAsFactors=FALSE, quote=""))
syn5_truth <- as.data.frame(read.table("/Users/cindyyu/Desktop/CS4220/Data/syn5/syn5_truth.bed",
                                header = FALSE, sep="\t",stringsAsFactors=FALSE, quote=""))
#might not need this
alltruth_df <- rbind(real1_truth, real2.1_truth, syn1_truth, syn2_truth, syn3_truth, syn4_truth, syn5_truth)

names(alltruth_df) <- c("Chr", "START_POS_REF", "END_POS_REF")

write.table(alltruth_df, "all_truths.bed",row.names = F,col.names = F, sep="\t", quote=FALSE)
```

#JOINS SETS SO THAT WE HAVE A TRUTH COL NEXT TO FEATURES
```{r}

#FOR JOINING TRUTH AND VCF FILES
jointruth<- function(truthdf, predsdf) {
  names(truthdf) <- c("Chr", "START_POS_REF", "END_POS_REF")
  truthdf$TRUTH = TRUE
  newdf <- predsdf %>% 
    left_join(truthdf, by=c('Chr','START_POS_REF', 'END_POS_REF'))
  newdf$TRUTH <- replace_na(newdf$TRUTH, FALSE)
  return(newdf)
}


#joins for vcfs and truths
df3 <- jointruth(alltruth_df, all_df)
real1_df <- jointruth(real1_truth, real1_df)
real2.1_df <- jointruth(real2.1_truth, real2.1_df)
syn1_df <- jointruth(syn1_truth, syn1_df)
syn2_df <- jointruth(syn2_truth, syn2_df)
syn3_df <- jointruth(syn3_truth, syn3_df)
syn4_df <- jointruth(syn4_truth, syn4_df)
syn5_df <- jointruth(syn5_truth, syn5_df)
```

#NUMBER OF POSITIONS AND PROPORTION OF VARIANTS
```{r}
df3 |>
  group_by(Sample_Name) |>
  tally()

counts <- df3 |>
  group_by(Sample_Name, TRUTH) |>
  tally()

ggplot(counts, aes(fill=TRUTH, y=n, x=Sample_Name)) + 
    geom_bar(stat="identity", width = 0.7)
```



#DATA IMPUTATION

##plotting, same group 
```{r}
#SHOWING WHERE TRUE FALSES ARE BY STACKED BAR PLOT
#sorry for the very inefficient code 

#MUTECT
df_m2na <- df3 |> 
  group_by(FILTER_Mutect2, is.na(m2_MQ)) |>
  tally() 

colnames(df_m2na)<- c("FILTER_Mutect2","Is_NA","n")

gg_m2na <- ggplot(df_m2na, aes(fill=Is_NA, y=n, x=FILTER_Mutect2)) + 
    geom_bar(stat="identity", width = 0.7)

#FREEBAYES
df_fbna <- df3 |> 
  group_by(FILTER_Freebayes, is.na(f_MQMR)) |>
  tally() 

colnames(df_fbna)<- c("FILTER_Freebayes","Is_NA","n")

gg_fbna <- ggplot(df_fbna, aes(fill=Is_NA, y=n, x=FILTER_Freebayes)) + 
    geom_bar(stat="identity", width = 0.7)

#VARDICT
df_vdna <- df3 |> 
  group_by(FILTER_Vardict, is.na(vd_SSF), is.na(vd_MSI)) |>
  tally() 

df_vdna <- df_vdna[-3]
colnames(df_vdna)<- c("FILTER_Vardict","Is_NA","n")

#MENTION THAT NA IN SSV ALSO HAD NA IN MSI
gg_vdna <- ggplot(df_vdna, aes(fill=Is_NA, y=n, x=FILTER_Vardict)) + 
    geom_bar(stat="identity", width = 0.7)

#VARSCAN
df_vsna <- df3 |> 
  group_by(FILTER_Varscan, is.na(vs_SPV), is.na(vs_SSC)) |>
  tally() 

df_vsna <- df_vsna[-3]
colnames(df_vsna)<- c("FILTER_Varscan","Is_NA","n")

#MENTION THAT NA IN SSV ALSO HAD NA IN MSI
gg_vsna <- ggplot(df_vsna, aes(fill=Is_NA, y=n, x=FILTER_Varscan)) + 
    geom_bar(stat="identity", width = 0.7)

df3 %>%
  ggplot(aes(x = FILTER_Mutect2, y = m2_MQ)) +
  geom_bar(stat = "identity")

```

#REPLACE W MEDIAN BY GROUP
```{r}

#replace NAs w MEAN
df3_narm <- df3 %>% mutate_all(~ifelse(is.na(.x), mean(.x, na.rm = TRUE), .x)) 

# #grouped replacing NAs w mean 
# df_groupmean <- df3 %>%
#   group_by(FILTER_Mutect2) %>%
#   mutate(m2_MQ = ifelse(is.na(m2_MQ), mean(m2_MQ, na.rm = TRUE), m2_MQ)) |>
#   ungroup() %>%
#   group_by(FILTER_Freebayes) %>%
#   mutate(f_MQMR = ifelse(is.na(f_MQMR), mean(f_MQMR, na.rm = TRUE), f_MQMR)) %>%
#   ungroup() %>%
#   group_by(FILTER_Vardict) %>%
#   mutate(vd_SSF = ifelse(is.na(vd_SSF), mean(vd_SSF, na.rm = TRUE), vd_SSF)) %>%
#   ungroup() %>%
#   group_by(FILTER_Vardict) %>%
#   mutate(vd_MSI = ifelse(is.na(vd_MSI), mean(vd_MSI, na.rm = TRUE), vd_MSI)) %>%
#   ungroup() %>%
#   group_by(FILTER_Varscan) %>%
#   mutate(vs_SSC = ifelse(is.na(vs_SSC), mean(vd_MSI, na.rm = TRUE), vs_SSC)) %>%
#   ungroup() %>%
#   group_by(FILTER_Varscan) %>%
#   mutate(vs_SPV = ifelse(is.na(vs_SPV), mean(vs_SPV, na.rm = TRUE), vs_SPV)) %>%
#   ungroup()

df_groupmedian <- df3 %>%
  group_by(FILTER_Mutect2) %>%
  mutate(m2_MQ = ifelse(is.na(m2_MQ), median(m2_MQ, na.rm = TRUE), m2_MQ)) |>
  ungroup() %>%
  group_by(FILTER_Freebayes) %>%
  mutate(f_MQMR = ifelse(is.na(f_MQMR), median(f_MQMR, na.rm = TRUE), f_MQMR)) %>%
  ungroup() %>%
  group_by(FILTER_Vardict) %>%
  mutate(vd_SSF = ifelse(is.na(vd_SSF), median(vd_SSF, na.rm = TRUE), vd_SSF)) %>%
  ungroup() %>%
  group_by(FILTER_Vardict) %>%
  mutate(vd_MSI = ifelse(is.na(vd_MSI), median(vd_MSI, na.rm = TRUE), vd_MSI)) %>%
  ungroup() %>%
  group_by(FILTER_Varscan) %>%
  mutate(vs_SSC = ifelse(is.na(vs_SSC), median(vd_MSI, na.rm = TRUE), vs_SSC)) %>%
  ungroup() %>%
  group_by(FILTER_Varscan) %>%
  mutate(vs_SPV = ifelse(is.na(vs_SPV), median(vs_SPV, na.rm = TRUE), vs_SPV)) %>%
  ungroup()


# df_groupmedian <- df3 %>%
#   group_by(FILTER_Mutect2, FILTER_Freebayes, FILTER_Vardict, FILTER_Varscan) %>%
#   summarize(median(m2_MQ))
# 
# df_groupmedian <- df3 %>%
#   group_by(FILTER_Freebayes) %>%
#   summarize(median(f_MQMR))
```

#MICE AGGR PLOT 
```{r}
md.pattern(df3)

df_hmm <- df3[, 9:19]

library(VIM)
aggr_plot <- aggr(df3, col=c('navyblue','red'), numbers=TRUE, sortVars=TRUE, labels=names(df3), cex.axis=.7, gap=3, ylab=c("Histogram of missing data","Pattern"))

```

#ACTUAL MICE IMPUTATION
```{r}
#missForest
library(missForest)

df3_features <- df3[13:19]
# df3_features$FILTER_Mutect2 = as.factor(df3_features$FILTER_Mutect2)
# df3_features$FILTER_Freebayes = as.factor(df3_features$FILTER_Freebayes)
# df3_features$FILTER_Vardict = as.factor(df3_features$FILTER_Vardict)
# df3_features$FILTER_Varscan = as.factor(df3_features$FILTER_Varscan)
df3_features$TRUTH = as.factor(df3_features$TRUTH)

library(mice)
#seed 10% missing values
df3.mis <- prodNA(df3_features, noNA = 0.1)
summary(df3.mis)

imputed_Data <- mice(df3.mis, m=5, maxit = 50, method = 'pmm', seed = 500)
summary(imputed_Data)


complete_data <- complete(imputed_Data, action = "long")

# Compare variance before and after imputation
var_original <- apply(df3, 2, var, na.rm = TRUE)
var_imputed <- apply(complete_data, 2, var)

# Scatter plot to compare
plot(var_original, var_imputed, main="Variance Comparison (Original vs. Imputed)")
abline(0,1, col="red") # Ideal case where variance is preserved

completeData1 <- complete(imputed_Data, 1)
completeData2 <- complete(imputed_Data, 2)
completeData3 <- complete(imputed_Data, 3)
completeData4 <- complete(imputed_Data, 4)
completeData5 <- complete(imputed_Data, 5)

write.table(completeData1, "impute_mice1.txt", row.names = F,col.names = F, sep=",", quote=FALSE)
write.table(completeData2, "impute_mice2.txt", row.names = F,col.names = F, sep=",", quote=FALSE)
write.table(completeData3, "impute_mice3.txt", row.names = F,col.names = F, sep=",", quote=FALSE)
write.table(completeData4, "impute_mice4.txt", row.names = F,col.names = F, sep=",", quote=FALSE)
write.table(completeData5, "impute_mice5.txt", row.names = F,col.names = F, sep=",", quote=FALSE)

read.table("impute_mice5.txt", header = F, sep = ',')
```


#MACHINE LEARNING RPART
```{r}
#remove chr and genomic start/stop

library(rpart)

#splitting
#SOMEHOW REMOVE COLUMNS OF GENOME NUMBER AND CHR
split_vals <- sample.split(df3_narm$TRUTH, SplitRatio = 0.80)
train_set <- subset(df3_narm, split_vals == TRUE)
test_set <- subset(df3_narm, split_vals == FALSE)

train_set_split <- train_set[,-(1:8)]

#model
rpart(as.factor(TRUTH)~ ., data = train_set_split) -> mod_class
predict(mod_class, test_set, type = 'class') -> result_class
table(test_set$TRUTH, result_class)

#refitting predictions to be bed file
#REALIZED I SHOULDNT HAVE UPENDED TO DF3, BUT RATHER TEST SET!
predictions <- data.frame(Chr = test_set$Chr,
                          START_POS_REF = test_set$START_POS_REF,
                          END_POS_REF = test_set$END_POS_REF, 
                          TRUTH=result_class)

predictions <- predictions |>
  filter(TRUTH == TRUE) 

my_predictions <- predictions[, -4]

write.table(my_predictions, "my-predictions.bed",row.names = F,col.names = F, sep="\t", quote=FALSE) 
```

#MACHINE LEARNING NARM
```{r}
df3_narm <- df3 %>% mutate_all(~ifelse(is.na(.x), mean(.x, na.rm = TRUE), .x)) 
 

#converting to factor!!!!!
df3_narm$TRUTH <- as.factor(df3_narm$TRUTH)

#split into test/train
set.seed(123)
split <- sample.split(df3_narm$TRUTH, SplitRatio = 0.8)

training_set <- subset(df3_narm, split == TRUE)
test_set <- subset(df3_narm, split == FALSE)

#training model
rf_model_mean <- randomForest(TRUTH ~ ., data = training_set, ntree = 100, importance = TRUE)

#predict for test set
test_predictions_mean <- predict(rf_model_mean, newdata = test_set)

confusion_matrix_mean <- table(test_set$TRUTH, test_predictions_mean)
print(confusion_matrix_mean) #F1: 0.9926


importance(rf_model_mean)
varImpPlot(rf_model) #unsure what this does


test_pred <- test_predictions_mean
# write.table(test_pred, "test_predictions_vector.bed",row.names = F,col.names = F, sep="\t", quote=FALSE)

#test_pred <-read.table("REALtest_predictions.bed", sep="\t" )


#PREDICTIONS INTO BED
test_setnew <- test_set[,1:3] #chromo and start/stops
test_setnew$TRUTH <- test_pred #new col w preds

#removes rows/cols so just variants
test_setnew <- test_setnew |>
  filter(TRUTH == TRUE)
test_setnew <- test_setnew[1:3]
  
write.table(test_set, "meantest_predictions.bed",row.names = F,col.names = F, sep="\t", quote=FALSE)
  
#TRUTH SET INTO BED
testset_truth <- test_set[, -(4:18)] |>
  filter(TRUTH == TRUE)
  
write.table(testset_truth, "meantest_TRUTH.bed",row.names = F,col.names = F, sep="\t", quote=FALSE)
```

```{r}

```


#MACHINE LEARNING SINGLE CASE IMPUTATION MEDIAN
```{r}
#processing + split into test/train
df_groupmedian$TRUTH <- as.factor(df_groupmedian$TRUTH)
set.seed(123)

split <- sample.split(df_groupmedian$TRUTH, SplitRatio = 0.8)
training_set <- subset(df_groupmedian, split == TRUE)
test_set <- subset(df_groupmedian, split == FALSE)


# Assign weights inversely proportional to sample size
sample_counts <- table(df_groupmedian$Sample_Name)
weights <- max(sample_counts) / sample_counts

# Train a weighted Random Forest
rf_model_median <- randomForest(
  TRUTH ~ ., 
  data = training_set, 
  ntree = 100, 
  strata = df_groupmedian$Sample_Name, 
  sampsize = rep(min(sample_counts), length(sample_counts)), 
  importance = TRUE
)


#actually training
training_featuresremoved <- training_set[, 13:20]
rf_model_median <- randomForest(TRUTH ~ ., data = training_featuresremoved, ntree = 100, importance = TRUE)
test_predictions_median <- predict(rf_model_median, newdata = test_set) #predicts
confusion_matrix <- table(test_set$TRUTH, test_predictions_median)
print(confusion_matrix)

importance(rf_model_median)
varImpPlot(rf_model) #unsure what this does


#PREDICTIONS INTO BED
test_setnew <- test_set[,1:3] 
test_setnew$TRUTH <- test_pred

test_setnew <- test_setnew |>
  filter(TRUTH == TRUE)

test_setnew <- test_setnew[1:3]

write.table(test_set, "mediantest_predictions.bed",row.names = F,col.names = F, sep="\t", quote=FALSE)

#TEST SET INTO BED
testset_truth <- test_set[, -(4:18)] |>
  filter(TRUTH == TRUE)


write.table(testset_truth, "mediantest_TRUTH",row.names = F,col.names = F, sep="\t", quote=FALSE)
```

```{r}
test_set_real1 <- test_set |>
  filter(Sample_Name == "icgc_cll-T") 
test_predictions_median <- predict(rf_model_median, newdata = test_set_real1) #predicts

confusion_matrix <- table(test_set_real1$TRUTH, test_predictions_median)
print(confusion_matrix)


test_set_notruth <- test_set[,-19]
test_predictions_median <- predict(rf_model_median, newdata = test_set_notruth) #predicts

confusion_matrix <- table(test_set$TRUTH, test_predictions_median)
print(confusion_matrix)


test_set_real1 <- test_set |>
  filter(Sample_Name == "dream1-T") 
test_predictions_median <- predict(rf_model_median, newdata = test_set_real1) #predicts

confusion_matrix <- table(test_set_real1$TRUTH, test_predictions_median)
print(confusion_matrix)

```



#MACHINE LEARNING MICE
```{r}

set.seed(123)

mice1 <- read.table('impute_mice1.txt', sep = ',', col.names = c("m2_MQ", "f_MQMR", "vs_SSC", "vs_SPV", "vd_SSF", "vs_MSI", "TRUTH"))

df_mice1 <- df3
df_mice1[13:19] <- mice1

df_mice1$TRUTH <- as.factor(df_mice1$TRUTH)
split <- sample.split(df_mice1$TRUTH, SplitRatio = 0.8)


training_set <- subset(df_mice1, split == TRUE)
test_set <- subset(df_mice1, split == FALSE)


rf_model_mice <- randomForest(TRUTH ~ ., data = training_set, ntree = 100, importance = TRUE)

test_predictions_mice <- predict(rf_model_mice, newdata = test_set_notruth)

confusion_matrix_mice <- table(test_set$TRUTH, test_predictions_mice)
print(confusion_matrix_mice)

importance(rf_model)
varImpPlot(rf_model)


test_pred <- test_predictions
write.table(test_pred, "test_predictions_vector.bed",row.names = F,col.names = F, sep="\t", quote=FALSE)

#test_pred <-read.table("REALtest_predictions.bed", sep="\t" )


#PREDICTIONS INTO BED
test_setnew <- test_set[,1:3] 
test_setnew$TRUTH <- test_pred

test_setnew <- test_setnew |>
  filter(TRUTH == TRUE)

test_setnew <- test_setnew[1:3]


write.table(test_set, "REALtest_predictions.bed",row.names = F,col.names = F, sep="\t", quote=FALSE)

#TEST SET INTO BED
testset_truth <- test_set[, -(4:18)] |>
  filter(TRUTH == TRUE)


write.table(testset_truth, "REALtest_TRUTH.bed",row.names = F,col.names = F, sep="\t", quote=FALSE)
```
#MODEL FOR MEDIAN IMPUTATION
```{r}
df3_median <- df3 %>% mutate_all(~ifelse(is.na(.x), median(.x, na.rm = TRUE), .x)) 
 
  #converting to factor!!!!!
df3_median$TRUTH <- as.factor(df3_median$TRUTH)
df3_median$Sample_Name <- as.factor(df3_median$Sample_Name)

set.seed(123)

split <- sample.split(df3_median$TRUTH, SplitRatio = 0.8)
training_set <- subset(df3_median, split == TRUE)
test_set <- subset(df3_median, split == FALSE)

rf_model_median <- randomForest(TRUTH ~ ., data = training_set, ntree = 100, importance = TRUE)

test_predictions_median <- predict(rf_model_median, newdata = test_set)

confusion_matrix_median <- table(test_set$TRUTH, test_predictions_median)
print(confusion_matrix_median) #F1: 0.9926

```

```{r}
library(ranger)
df3_median$TRUTH <- as.factor(df3_median$TRUTH)
df3_median$Sample_Name <- as.factor(df3_median$Sample_Name)


# Compute sample-level weights (inverse frequency)
sample_counts <- table(df3_median$Sample_Name)
sample_weights <- max(sample_counts) / sample_counts

# Compute mutation type-level weights
mutation_counts <- table(df3_median$TRUTH)
mutation_weights <- max(mutation_counts) / mutation_counts

# Assign combined case weights
names(case_weights) <- levels(df3_median$Sample_Name)
case_weights <- sample_weights* mutation_weights


case_weights <- as.data.frame(case_weights)
names(case_weights) <- c('Sample_Name', 'Freq')

case_weights2 <- case_weights
case_weights2[6,2] <- 2 #WITH DIFFERENT WEIGHTING


df3_median <- df3_median %>%
  left_join(case_weights, by = 'Sample_Name')

split <- sample.split(df3_median$TRUTH, SplitRatio = 0.8)
training_set <- subset(df3_median, split == TRUE)
test_set <- subset(df3_median, split == FALSE)

# Train the weighted Random Forest model
rf_model_median <- ranger(
  TRUTH ~ ., 
  data = training_setfeatures, #remember to change this to training lol
  case.weights = training_set$Freq, 
  num.trees = 100
)

print(rf_model)

test_set_notruth <- test_set[,-19] #removes truth(prob unnecessary)

predictions <- predict(rf_model, data = test_set_notruth)

predicted_classes <- predictions$predictions
print(predicted_classes)  # Displays predicted classes
length(test_set_notruth$TRUTH)

confusion_matrix_median <- table(test_set$TRUTH, predicted_classes)
print(confusion_matrix_median) #F1: 0.9926
```


#IDK HOW TO GET THIS WORKING BUT WE GOTTA GET IT
```{r}
# install.packages("pROC")
# library(pROC)
# 
# length(test_set)
# 
# roc_curve <- roc(test_set$TRUTH, test_predictions_median[, 4])
# 
# #plott!!
# plot(roc_curve, main = "ROC Curve", col = "blue", lwd = 2)
# 
# 
# auc_value <- auc(roc_curve)
# print(paste("AUC: ", auc_value))
# 
# 
# confusion_matrix <- table(test_set$correct, predict(rf_model, newdata = test_set))
# print(confusion_matrix)

```

=
```{r}
importance(rf_model_mice)
```



grid search 



library(rpart)
library(caTools)
library(tidyr)

# Define the grid of hyperparameters to search
param_grid <- expand.grid(
  cp = c(0.01, 0.05, 0.1, 0.2),        # Complexity parameter
  maxdepth = c(3, 5, 7, 10),           # Maximum depth of tree
  minsplit = c(20, 30, 40),            # Minimum number of observations required for a split
  stringsAsFactors = FALSE
)

# Create empty list to store results
results <- list()

# Set up the data
split_vals <- sample.split(df3_narm$TRUTH, SplitRatio = 0.80)
train_set <- subset(df3_narm, split_vals == TRUE)
test_set <- subset(df3_narm, split_vals == FALSE)

train_set_split <- train_set[, -(1:8)]  # Removing unnecessary columns

# Perform Grid Search
for (i in 1:nrow(param_grid)) {
  
  # Extract hyperparameters from the grid
  cp_value <- param_grid$cp[i]
  maxdepth_value <- param_grid$maxdepth[i]
  minsplit_value <- param_grid$minsplit[i]
  
  # Train the decision tree model with the current hyperparameters
  mod_class <- rpart(as.factor(TRUTH) ~ ., 
                     data = train_set_split,
                     cp = cp_value,
                     maxdepth = maxdepth_value,
                     minsplit = minsplit_value)
  
  # Make predictions on the test set
  result_class <- predict(mod_class, test_set, type = 'class')
  
  # Calculate performance metric (accuracy in this case)
  accuracy <- sum(test_set$TRUTH == result_class) / nrow(test_set)
  
  # Store the result
  results[[i]] <- list(
    cp = cp_value,
    maxdepth = maxdepth_value,
    minsplit = minsplit_value,
    accuracy = accuracy
  )
}

# Convert results to a data frame
results_df <- do.call(rbind, lapply(results, as.data.frame))

# Find the best hyperparameters (highest accuracy)
best_model <- results_df[which.max(results_df$accuracy), ]
print(best_model)

# Optional: Use the best hyperparameters to refit the model
best_mod_class <- rpart(as.factor(TRUTH) ~ ., 
                        data = train_set_split,
                        cp = best_model$cp,
                        maxdepth = best_model$maxdepth,
                        minsplit = best_model$minsplit)

# Make predictions with the best model
final_result_class <- predict(best_mod_class, test_set, type = 'class')

# Display confusion matrix for the best model
table(test_set$TRUTH, final_result_class)





lol you might have this already? 



```{r}

# Load necessary libraries
library(rpart)
library(caTools)

# Split the data
split_vals <- sample.split(df3_narm$TRUTH, SplitRatio = 0.80)
train_set <- subset(df3_narm, split_vals == TRUE)
test_set <- subset(df3_narm, split_vals == FALSE)

# Remove unnecessary columns
train_set_split <- train_set[, -(1:8)]

# Create a dataframe to store the results of each model's performance
results_df <- data.frame(cp = numeric(), maxdepth = numeric(), minsplit = numeric(), accuracy = numeric())

# Loop over each combination of hyperparameters
for(i in 1:nrow(param_grid)) {
  # Set the current hyperparameters
  cp_val <- param_grid$cp[i]
  maxdepth_val <- param_grid$maxdepth[i]
  minsplit_val <- param_grid$minsplit[i]
  
  # Train the model with the current hyperparameters
  mod_class <- rpart(as.factor(TRUTH) ~ ., data = train_set_split, 
                     control = rpart.control(cp = cp_val, maxdepth = maxdepth_val, minsplit = minsplit_val))
  
  # Make predictions
  result_class <- predict(mod_class, test_set, type = 'class')
  
  # Calculate accuracy
  accuracy <- sum(result_class == test_set$TRUTH) / nrow(test_set)
  
  # Save the results
  results_df <- rbind(results_df, data.frame(cp = cp_val, maxdepth = maxdepth_val, minsplit = minsplit_val, accuracy = accuracy))
}

# Print the results of the grid search
print(results_df)






```


finding the best model:

```{r}


# Get the best model based on accuracy
best_model <- results_df[which.max(results_df$accuracy), ]
print(best_model)




```



